---
title: "Applied Survey Research: Questionnaire Development and Pilot Evaluation"
author: "Monika Bagociute"
date: "`r format(Sys.time(), '%B %d, %Y')`"
output:
  html_document:
    toc: true
    toc_float: true
    code_folding: hide
---


## Research overview

This case study documents an applied quantitative survey research workflow,
from questionnaire development and pilot testing to quantitative evaluation.

The analysis demonstrates how pilot data can be used to assess data quality,
questionnaire functioning, item-level patterns, internal consistency, and
differences between predefined participant groups.

### Research objectives

The pilot analysis focuses on:

- preparing and harmonising questionnaire data;
- evaluating item-level response patterns;
- assessing internal consistency of questionnaire scales;
- comparing predefined participant groups;
- identifying issues that may inform questionnaire refinement before the main study.

### Analytical workflow

The analysis follows four main stages:

1. Data preparation and questionnaire harmonisation
2. Descriptive and item-level analysis
3. Group comparisons
4. Reliability and correlation analysis

### Data note

This portfolio version is a sanitised representation of an applied research
project conducted in a professional research setting. Confidential research
materials, identifiable participant information, and original source files are
not included in the public portfolio repository.


```{r setup, message=FALSE, warning=FALSE}
library(readxl)
library(dplyr)
library(stringr)
library(psych)
library(ggplot2)
library(tidyr)
library(knitr)
```

### 1. Load data 

```{r}

# The original participant-level dataset is confidential and is not included
# in the public portfolio repository.
#
# The following code generates a synthetic dataset with the same analytical
# structure as the pilot data.

set.seed(2026)

n <- 69
n_items <- 34

data <- data.frame(
  participant = 1:n,
  tto = c(rep(0, 30), rep(1, 39))
)

# Generate questionnaire responses on a 1–5 scale.
# Items are simulated with moderate correlations within conceptual scales.

simulate_scale <- function(n, n_items, mean = 3.5, sd = 0.7) {
  latent <- rnorm(n, mean = mean, sd = sd)
  
  responses <- sapply(
    1:n_items,
    function(i) {
      round(
        pmin(
          5,
          pmax(
            1,
            latent + rnorm(n, 0, 0.6)
          )
        )
      )
    }
  )
  
  as.data.frame(responses)
}

PD <- simulate_scale(n, 11, mean = 3.5)
Language <- simulate_scale(n, 14, mean = 3.7)
GC <- simulate_scale(n, 9, mean = 3.6)

data <- cbind(
  data,
  PD,
  Language,
  GC
)

names(data)[3:36] <- paste0("Q", 1:34)

Qvars <- paste0("Q", 1:34)

# Add a modest synthetic group difference to selected items.
data[data$tto == 1, c("Q12", "Q13", "Q14", "Q15")] <-
  pmin(5, data[data$tto == 1, c("Q12", "Q13", "Q14", "Q15")] + 0.3)




```

### 2. Data cleaning


The original pilot dataset required several preparation steps before analysis.
These included:

- filtering responses according to the study's participation criteria;
- harmonising questionnaire variables;
- handling structural missingness created by questionnaire branching;
- combining equivalent variables across questionnaire language versions;
- standardising response coding;
- converting questionnaire responses to numeric analysis variables.

The public portfolio version uses synthetic data and therefore does not reproduce
participant-level cleaning decisions from the original research dataset.


#### Harmonise questionnaire language versions

The pilot included questionnaire versions in different languages. Equivalent
items were mapped to a common set of questionnaire variables so that responses
could be analysed consistently across participants.

For each questionnaire item, responses from the relevant language-specific
variables were combined into a single analysis variable. Where one version was
not populated because of the questionnaire routing, the corresponding response
from the other language version was used.

This harmonisation step created a common set of 34 questionnaire items for the
subsequent analyses.



#### Standardise response coding

Questionnaire responses were originally collected using categorical response
labels. Because the questionnaire included different language versions,
equivalent response categories were mapped to a common numeric scale.

The harmonised items were coded from 1 to 5, preserving the ordinal structure
of the response options.

The public portfolio version uses anonymised synthetic response categories.




### 3. Sample description

The synthetic dataset contains 69 observations and represents the structure of
the pilot sample used in the original analysis. Participants are divided into
two predefined groups for the purpose of demonstrating the comparative
analysis.

```{r}
kable(data.frame(
  Participants=nrow(data),
  TTO=sum(data$tto==1),
  Non_TTO=sum(data$tto==0)
))
```




### 4. Descriptive statistics

Descriptive statistics were used to examine the distribution and central
tendency of responses across questionnaire items. Item-level means provide an
initial overview of response patterns and help identify items that may warrant
closer examination during pilot evaluation.

```{r}
describe(data[Qvars])  # Q4 and Q16 
```

#### Mean response by questionnaire item

```{r}
means <- colMeans(data[Qvars], na.rm = TRUE)

plotdf <- data.frame(
  Question = Qvars,
  Mean = means
)

ggplot(
  plotdf,
  aes(reorder(Question, Mean), Mean)
) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  labs(
    x = NULL,
    y = "Mean response"
  )



```
### 5. Comparison between predefined groups

The pilot analysis compared questionnaire responses between two predefined
participant groups. Independent-samples t-tests were used to examine
item-level differences in mean responses.

These comparisons are exploratory and are intended to identify patterns that
may warrant further investigation in the main study.

```{r}
results <- data.frame()

for(i in 1:34){

q <- paste0("Q",i)

test <- t.test(data[[q]]~data$tto)

means <- tapply(data[[q]],data$tto,mean)

results <- rbind(results,
data.frame(
Question=q,
Mean_nonTTO=means["0"],
Mean_TTO=means["1"],
t=test$statistic,
p=test$p.value))
}

kable(results,digits=3)
```



#### Mean difference by questionnaire item
```{r}
plot_results <- results |>
  mutate(
    Difference = Mean_TTO - Mean_nonTTO
  )

ggplot(
  plot_results,
  aes(
    reorder(Question, Difference),
    Difference
  )
) +
  geom_col() +
  coord_flip() +
  theme_minimal() +
  labs(
    x = NULL,
    y = "Mean difference between groups"
  )
```



- It is clear that the language-related questions distinguish the two groups the most.


```{r}
# Adjust p-values for multiple comparisons
results$p_adjusted <- p.adjust(
  results$p,
  method = "BH"
)

kable(
  results,
  digits = 3,
  caption = "Item-level group comparisons with Benjamini-Hochberg adjusted p-values"
)

```

Because the analysis evaluates multiple questionnaire items separately,
unadjusted p-values can increase the risk of false-positive findings. The
Benjamini-Hochberg procedure is therefore used to control the false discovery
rate across the item-level comparisons.

The pilot comparisons are exploratory and should not be interpreted as
confirmatory evidence of group differences.


### 6. Internal consistency

Internal consistency was evaluated separately for the three questionnaire
scales using Cronbach's alpha and item-level diagnostics.

The purpose of this analysis was to assess whether items within each proposed
scale showed a coherent pattern of responses and to identify potential
measurement issues for consideration before the main study.



The analysis reports Cronbach's alpha together with item-level diagnostics.
These statistics are interpreted alongside the item correlations rather than
as a standalone measure of scale quality.

#### Personal Development
```{r}

PD <- psych::alpha(data[,paste0("Q",1:11)])
PD
```


#### Language

```{r}
Language <- psych::alpha(data[,paste0("Q",12:25)])
Language
```



#### Global Citizenship

```{r}
GC <- psych::alpha(data[,paste0("Q",26:34)])
GC
```


#### Reliability comparison

```{r}
alphas <- data.frame(
Scale=c("Personal Development",
"Language",
"Global Citizenship"),
Alpha=c(PD$total$raw_alpha,
Language$total$raw_alpha,
GC$total$raw_alpha))

ggplot(
  alphas,
  aes(
    x = Scale,
    y = Alpha
  )
) +
  geom_col() +
  geom_hline(
    yintercept = 0.70,
    linetype = "dashed"
  ) +
  coord_cartesian(ylim = c(0, 1)) +
  theme_minimal() +
  labs(
    x = NULL,
    y = "Cronbach's alpha"
  )
```

The reliability estimates provide an initial assessment of internal
consistency across the proposed scales. In a full measurement evaluation,
these results would be considered alongside item content, inter-item
correlations, dimensionality, and the intended use of the scale.

### 7. Item-level correlation structure

Examining the correlation matrix provides an additional view of the
relationships among questionnaire items.

This analysis can help assess whether items intended to represent related
constructs show coherent response patterns and whether some items appear
unusually weakly or strongly related to the rest of their scale.


```{r}
corr <- cor(data[Qvars],use="pairwise.complete.obs")

heatmap(corr,scale="none")
```

## Reproducibility

The public analysis uses synthetic data generated within the R Markdown
document. A fixed random seed is used so that the synthetic dataset and
resulting analyses can be reproduced.

The original participant-level research data are not included in this
portfolio.


